R for Clinical Study Reports and Submission

### 2021-10-04 project-folder.Rmd
fix typo

The `man` folder is used to save documentation automatically generated by `roxygen2`.
A typical workflow is to add `roxygen2` documentations at the header of each function in the `R\` folder. 
Then `devtools::documents()` is used to generate all the documentation files in the `man\` folder. 
More details can be found in [Chapter 10 of the R packages book](https://r-pkgs.org/man.html).

@@

Reproducibility of analysis is one of the most important aspects for regulatory deliverables. 
By using R package folder structure and proper tools (e.g. `renv`, `packrat`), 
we illustrate how to achieve reproducibility for R and R package versions.

> This is the same level of reproducibility in most SAS environment.
> https://support.sas.com/techsup/pcn/altopsys.html

@@

We can also define the package repository to be a specific snapshot date.
For example, we used Microsoft R Application Network (MRAN) to 
define the snapshot date to be `2021-08-06`. 
The snapshot date freeze the R package repository.

@@

> RStudio Package Manager (RSPM) provided a solution to host both 
> public avaiable and internally developed R packages. 
> However, public RSPM server (https://packagemanager.rstudio.com/client/#/repos/2/overview) 
> did not provide daily snapshot as MRAN had.

Secondly, we use `renv` to lock R package versions and saved in `renv.lock` file. 
The `renv` provide a robust and stable approach to manage R package version 
for a project specific R packages.
An introduction of `renv` can be found in its 
[package website](https://rstudio.github.io/renv/articles/renv.html).

@@

> A cloud based R environment (e.g. RStudio Workbench) 
> can enhance the reproducibility within an organization 
> by using the same operation system, R and R package versions for an A&R project. 
> More details can be found in [https://environments.rstudio.com/](https://environments.rstudio.com/).

@@

- [`devtools`](https://devtools.r-lib.org/): make package development easier. 
  + A good overview can be found in [Chapter 2 of the R packages](https://r-pkgs.org/whole-game.html) book  
  + `devttols::load_all()`: load all functions in `R/` folder and running environment. 
  + `devtools::document()`: automatically create documentation using `Roxygen2`
  + `devttols::check()`: automatically perform compliance check as an R package. 
  + `devtools::build_site()`: automatically run analysis scripts in batch and create a `pkgdown` website. 
- [`usethis`](https://usethis.r-lib.org/): automates repetitive tasks that arise during project setup and development.
- [`testthat`](https://testthat.r-lib.org/): streamline testing code.
  + A discussion of using the `testthat` for an A&R project can be found in (@madhutest).  
- [`pkgdown`](https://pkgdown.r-lib.org/): generate static html documentation for an R package
  + It also allow you to run all analysis code in batch.

@@

As the consistent folder is defined, it also become easier to create specific tools 
that fit for the analysis and reporting purpose.

@@

- Create project template using [RStudio Project Templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html)
- Add additional compliance check for analysis and reporting. 
- Save log files for running in batch.

>>>>

The `man` folder is used to save documentation automatically generated by `roxygen2`.
A typical workflow is to add `roxygen2` documentation at the header of each function in the `R\` folder. 
Then `devtools::documents()` is used to generate all the documentation files in the `man\` folder. 
More details can be found in [Chapter 10 of the R packages book](https://r-pkgs.org/man.html).

@@

Reproducibility of analysis is one of the most important aspects of regulatory deliverables. 
By using the R package folder structure and proper tools (e.g. `renv`, `packrat`), 
we illustrate how to achieve reproducibility for R and R package versions.

> This is the same level of reproducibility in most SAS environments.
> https://support.sas.com/techsup/pcn/altopsys.html

@@

We can also define the package repository to be a specific snapshot date.
For example, we used Microsoft R Application Network (MRAN) to 
define the snapshot date to be `2021-08-06`. 
The snapshot date freezes the R package repository.

@@

> RStudio Package Manager (RSPM) provided a solution to host both 
> publicly available and internally developed R packages. 
> However, the public RSPM server (https://packagemanager.rstudio.com/client/#/repos/2/overview) 
> did not provide a daily snapshot as MRAN had.

Secondly, we use `renv` to lock R package versions and saved in `renv.lock` file. 
The `renv` provides a robust and stable approach to manage R package versions 
for a project specific R packages.
An introduction of `renv` can be found on its 
[package website](https://rstudio.github.io/renv/articles/renv.html).

@@

> A cloud-based R environment (e.g. RStudio Workbench) 
> can enhance the reproducibility within an organization 
> by using the same operation system, R and R package versions for an A&R project. 
> More details can be found in [https://environments.rstudio.com/](https://environments.rstudio.com/).

@@

- [`devtools`](https://devtools.r-lib.org/): make package development easier. 
  + A good overview can be found in [Chapter 2 of the R packages](https://r-pkgs.org/whole-game.html) book  
  + `devttols::load_all()`: load all functions in `R/` folder and running environment. 
  + `devtools::document()`: automatically create documentation using `Roxygen2`
  + `devttols::check()`: automatically perform compliance check as an R package. 
  + `devtools::build_site()`: automatically run analysis scripts in batch and create a `pkgdown` website. 
- [`usethis`](https://usethis.r-lib.org/): automates repetitive tasks that arise during project setup and development.
- [`testthat`](https://testthat.r-lib.org/): streamline testing code.
  + A discussion of using the `testthat` for an A&R project can be found in (@madhutest).  
- [`pkgdown`](https://pkgdown.r-lib.org/): generate static html documentation for an R package
  + It also allows you to run all analysis code in batch.

@@

As the consistent folder is defined, it also becomes easier to create specific tools 
that fit for the analysis and reporting purpose.

@@

- Create project template using [RStudio Project Templates](https://rstudio.github.io/rstudio-extensions/rstudio_project_templates.html)
- Add additional compliance checks for analysis and reporting. 
- Save log files for running in batch.
